### 2024-03-19 04b_supplementary_plots.Rmd
SI plots

---
title: "supplementary plots"
output: html_document
date: "2024-02-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(arrow)
library(lubridate)
library(tidytext)
library(ggpattern)
library(ggpubr)
library(ggh4x)
```

## Read in the data and model metadata
```{r read-in-data}
targets <- readr::read_csv("https://data.ecoforecast.org/neon4cast-targets/aquatics/aquatics-targets.csv.gz") |> 
  na.omit()

lake_sites <- c('TOOK', 'PRLA','PRPO', 'CRAM', 'LIRO','BARC', 'SUGG')
cols_modeltype <- c('empirical' = "#482576FF",
                    'MME' =  "#35608DFF",
                    'process' = "#21908CFF",
                    'ML' = '#BBDF27FF',
                    'null' = '#A9A9A9')
# read in the metadata from the google sheet
googlesheets4::gs4_deauth()
model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")

# Open dataset-------------
# start with lake temperatures
temperature_scores <- arrow::open_dataset("./scores_2023") |>
  dplyr::filter(!(model_id %in% c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline') &
                  site_id == 'TOOK')) |> 
  filter(site_id %in% lake_sites,
         model_id %in% model_meta$model_id,
         variable == 'temperature') |>
  collect()

# Calculate skill scores ------------
temperature_climatology <- arrow::open_dataset("./scores_2023") |>
  filter(model_id == 'climatology', 
         site_id %in% lake_sites,
         variable == 'temperature') |>
  collect() |> 
  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
  arrange(reference_datetime, datetime, site_id) |> 
  pivot_wider(values_from = c(crps, logs),
              names_from = model_id)

temperature_skill <- temperature_scores |>
  filter(model_id != 'climatology') |> 
  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
  full_join(temperature_climatology) |>

  # calcaulte the skill relative to climatology (- is bad, + is good)
  mutate(skill_crps = crps_climatology - crps,
         skill_logs = logs_climatology - logs,
         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  # consistent forecast period
  filter(horizon <= 30,
         horizon >= 1, 
         horizon != 15) |> 
  arrange(reference_datetime, site_id, datetime, model_id) 
```
## Figure s1 - targets timeseries

```{r}
targets_ts <- targets |> 
  filter(year(datetime) < 2023) |> 
  ggplot(aes(x=datetime, y=variable, colour = variable)) +
  geom_point(show.legend = F) +
  facet_wrap(~site_id, ncol = 5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave(plot = targets_ts, 'Plots/FigureS1.png', width = 20, height = 15, units = 'cm')
```

## Figure S2 - all models, skill over horizon

```{r}
horizon_skill <- 
  temperature_skill |> 
  filter(!(model_id %in% c(top_mods$model_id)), 
         horizon > 0, 
         horizon != 15,
         horizon <=30) |> 
  group_by(model_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  left_join(select(model_meta, model_id, manuscript_name)) |> 
  # mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name))) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = manuscript_name, linetype = manuscript_name)) +
  geom_hline(yintercept = 0, colour = 'grey40' ) + 
  geom_line(show.legend = F)  +  
  labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +
  theme_bw()

ggsave(plot = horizon_skill, 'Plots/FigureS2.png', width = 10, height = 10, units = 'cm')
```

## Figure S3 - performance of null models

```{r}
baseline_forecasts <- arrow::open_dataset("scores_2023") |>
  filter(site_id %in% lake_sites,
         model_id %in% c('persistenceRW', 'climatology')) |>
  collect()

baseline_forecasts |> 
  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  group_by(site_id, model_id, horizon) |> 
  summarise(crps_median = median(crps, na.rm = T), .groups = 'drop') |> 
  ggplot(aes(x=horizon, y=crps_median, colour = site_id)) +
  geom_line() +
  facet_wrap(~model_id)

baseline_forecasts |> 
  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  group_by(site_id, model_id, reference_datetime) |> 
  summarise(crps_median = median(crps, na.rm = T), .groups = 'drop') |> 
  ggplot(aes(x = as_date(reference_datetime), y=crps_median, colour = site_id)) +
  geom_line() +
  facet_wrap(~model_id, scale = 'free_y')

```

## Figure S4

Plotting only dates where all sites are evaluated shows the same pattern as when all dates are considered. 
```{r overlapping-periods}
n_model_type <- temperature_skill |> 
  distinct(model_id) |> 
  left_join(model_meta) |>
  group_by(model_type) |> 
  summarise(n = n())  |> 
  mutate(site_id = 'TOOK', mean_skill = -3.25) |> 
  filter(model_type != 'null')

# top 10 average models
top_mods <- temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  slice_max(mean_crps, n = 10, na_rm = T) |> 
  distinct(model_id)

# what duration have the same dates, including TOOK
date_same_A <-  temperature_skill |> 
  na.omit() |> 
  group_by(site_id) |> 
  summarise(first_date = min(datetime), 
            last_date = max(datetime)) |> 
  summarise(first_date = max(first_date),
            last_date = min(last_date))

# what duration have the same dates, including TOOK
date_same_B <- temperature_skill |> 
  filter(site_id != 'TOOK') |> 
  na.omit()  |> 
  group_by(site_id) |> 
  summarise(first_date = min(datetime), 
            last_date = max(datetime)) |> 
  summarise(first_date = max(first_date),
            last_date = min(last_date))

A <- temperature_skill |> 
  left_join(model_meta, by = 'model_id') |>
  filter(model_type != 'null', 
         between(datetime, date_same_B$first_date, date_same_B$last_date)) |> 
  group_by(model_type, model_id, site_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
  ggplot(aes(x = mean_skill, 
             y = site_id, 
             colour = model_type)) +
  geom_vline(xintercept = 0, linetype = 'solid') +
  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
  geom_boxplot() +
  facet_wrap(~model_type, scales = 'free_y',nrow = 2) +
  labs(y='Site', x = 'mean relative skill (°C)') +
  scale_colour_manual(values = cols_modeltype, name = 'Model Type') +
  theme_bw()

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id, 
         between(datetime, date_same_B$first_date, date_same_B$last_date)) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = site_id)) +
  geom_hline(yintercept = 0, colour = 'grey40') + 
  geom_line()  +  
  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
  scale_color_viridis_d(option = 'H', name  = 'Site') +
  theme_bw()

cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'lr', labels = 'auto') |> 
  ggsave(filename = './Plots/FigureS4.png', width = 19, height = 16, units = 'cm')
```